Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of TESS

### 2023-07-14 paper/ms.tex
update ms

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f2.pdf}
    \vspace{-0.3cm}
		\caption{
      {\bf CQVs found in the TESS 2-minute data.}
      Phased TESS light curves over one month are shown for \ngoods\
      CQVs.  Gray are raw 2-minute data; black bins to 300 points per
      cycle.  Each panel is labeled by the TIC identifier, the TESS
      sector number, and the period in hours.  Objects are ordered
      such that sources with the most TESS data available are on top
      (see Section~\ref{sec:catalog}).
		}
		\label{fig:cqvs}
	\end{center}
\end{figure*}

@@

The upper age limit for the CQV phenomenon may even reach 200\,Myr,
based on the candidate membership of TIC~294328885 in the Carina Near
moving group \citep{2006ApJ...649L.115Z}.  This group's age primarily
rests on the lithium sequence of its G-dwarfs
\citep{2006ApJ...649L.115Z}, which shows a coeval population
of stars older than the Pleiades and younger than the 400\,Myr Ursa
Major moving group.  It would interesting to improve the
precision of the group's age measurement, for instance using
gyrochronology of F, G or K members, in order to verify this point.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f3.pdf}
		\vspace{-0.6cm}
		\caption{
		{\bf CQVs keep their periods but change their shapes.}
      32 of our \ngoods\ CQVs from Figure~\ref{fig:cqvs} had
      120-second cadence TESS data available for a baseline of at
      least two years; the 27 brightest are shown here.  Each panel
      shows one sector of TESS data, and is phased to its deepest
      minimum in flux.  Each panel's title shows the TIC identifier
      and approximate period in hours.  Text insets show the TESS
      sector numbers, which generally span two years, or at least
      1{,}000 cycles.  The vertical scale is fixed across sectors to
      clarify shape changes.
      Gray circles are raw 2-minute data; colored circles bin to 300
      points per cycle. 
		}
		\label{fig:evoln}
	\end{center}
\end{figure*}

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f4.pdf}
		\vspace{-0.35cm}
		\caption{
      {\bf LP 12-502 (TIC~402980664) light curve}, where each time
      chunk represents one TESS orbit.  Data were acquired in Sectors
      18-19, 25-26, 53, and 57-58.  Flares are drawn in gray.  The red
      vertical lines highlight apparently instantaneous state changes
      in the shape of the dip pattern.  The light curve is binned to
      15-minute intervals so that there are 96 points per day, and each
      point is connected by a line.  
      Data gaps longer than 15-minutes are not
      interpolated; if data are missing, nothing is plotted.
		}
		\label{fig:lplc}
	\end{center}
\end{figure*}

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f5.pdf}
		\vspace{-0.6cm}
		\caption{
      {\bf Evolution of LP 12-502} ($P$=18.6\,h) at fixed period and
      epoch over three years.  Each panel shows one (stacked) TESS
      orbit; small text denotes relative cycle number.  There are 200
      binned black points per cycle.  The TESS pointing law dictates
      time gaps; larger gaps tend to yield larger shape changes.  The
      dips usually evolve over tens to hundreds of cycles.  However
      cycles 1233-1264 show a dip that ``switched'' from a depth and
      duration of 3\% and 3\,hr to 0.3\% and 1\,hr over less than one
      cycle (cf.~Figure~\ref{fig:lplc}).
		}
		\label{fig:lp}
	\end{center}
\end{figure*}

@@

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=0.9\textwidth]{f6.pdf}
		\vspace{-0.1cm}
		\caption{
      {\bf Properties of CQVs identified by our search}.
      The \ngoods\ bonafide CQVs in Table~\ref{tab:thetable} are 
      the dark blue circles; \nmaybes\ candidate CQVs are light blue circles.
      The top panels show the \nstarssearched\ target stars with 120-second cadence TESS data
      as the shaded gray background;
      darker regions correspond to a larger relative number of searched stars.
      The lower-left panel compares the rotation--color distribution of CQVs against the rotation
      periods of K- and M-dwarfs in the Pleiades from \citet{2016AJ....152..114R}.
      The lower-middle panel plots the derived corotation radii
      $R_{\rm cr} = (GM/\Omega^2)^{1/3}$ in units of stellar radii against the measured
      CQV periods, in units of hours.
		}
		\label{fig:catalogscatter}
	\end{center}
\end{figure*}

@@

\begin{enumerate}[leftmargin=*]
	%
  \item CQVs are quasiperiodic.  The mean periods 
    remain fixed over the $>$1{,}000-cycle baseline of available
    observations; but the light curve shapes always evolve (Figure~\ref{fig:evoln}).
  %
  \item The same CQV can show dips with similar but clearly
    distinct periods.  LP 12-502, for instance, showed dips with four
    distinct periods within $\pm 0.3\%$ of its fundamental period,
    sometimes simultaneously, and each lasting for up to 50 cycles
    (Figure~\ref{fig:lpriver0}).
  %
    \item CQVs evolve over timescales that are both secular
    ($>$100\,cycles) and impulsive ($<$1 cycle).  ``State-switches''
    can cause dips to collapse instantaneously, and are often (but not
    exclusively) linked with observed optical flares.  Dip growth
    however seems to happen over durations of at least ten cycles, and
    slow dip decay can also occur.
  %
    \item The rate of dip evolution can be used to place a model-dependent
    constraint on amount of material that is either being accreted or ejected
    during the state changes (Section~\ref{subsec:massflux}).
    Order of magnitude estimates require at least an asteroid belt's worth
    of dust ($10^{-4}$\,$M_\oplus$) over $10^8$ years, or 100 times more 
    material if the occulting matter is gas.
  %
  \item The CQV phenomenon persists for $\gtrsim$150\,Myr, based on
    the existence of multiple CQVs in AB Dor, the Pleiades, and
    Psc-Eri (Section~\ref{subsec:starprops}).  It may even extend to
    200\,Myr, based on the one CQV we found in the Carina Near
    moving group (TIC~294328885; $\approx$200\,Myr).  The lack of
    detected CQVs in the Hyades and Praesepe suggests that the
    lifetime of the phenomenon is likely limited to the first few
    hundred million years.
	%
    \item The duty cycle for CQVs seems to be $\approx 75$\%, based on
    the fraction of bonafide CQVs that turned on or turned off during
    TESS re-observations, two years after their initial observation.
  %
    \item While the majority of detected CQVs have masses of
    0.1-0.5\,M$_\odot$, the upper mass limit for the CQV phenomenon
    might extend into the K-dwarf regime ($<$$0.8$\,M$_\odot$).
    TIC~405754448 and TIC~405910546, with masses of $\approx$0.8 and
    $\approx$0.6 solar masses, are the two massive outliers
    (Section~\ref{subsec:extreme}), and we encourage additional
    scrutiny of these objects in future work.
  %
    \item The closest CQVs to the Sun are at distances of 15-20\,pc, and
    the brightest CQVs have $V$$\approx$12 ($J$$\approx$7.5).  This
    conclusion is an extrapolation from the CQV sample derived in this work
    being $\gtrsim$80\% complete within 20\,pc.  Expanding our analysis of
    the TESS data to the full frame images would yield a truly
    volume-limited selection function, and would expand the CQV census
    by about a factor of two within 50\,pc, and by a factor of ten
    within 100\,pc.
\end{enumerate}

@@

\acknowledgments
This work was supported by the 
Heising-Simons 51~Pegasi~b Fellowship (LGB).
TIC 402980664 was observed at 120-second cadence thanks to the TESS Guest
Investigator programs G022252 (PI: J.~Schlieder; Sectors 18, 19, 25,
26) and G04168 (PI: R.~Jayaraman; Sector 53).
While many conversations helped encourage this work,
a few thank-yous in particular (from LGB) 
are owed to G.~Laughlin, A.~Mann, L.~Rebull, and J.~Spake for clarifications and ideas.

@@

\startlongtable
\begin{deluxetable}{rrrrrrlrllrrrr}
\tabletypesize{\footnotesize}
%\tabletypesize{\scriptsize}
\tablecaption{Complex quasiperiodic variables identified in the TESS 2-minute data. \label{tab:thetable}}
%\toprule
%TIC & $T$ & $d$ & BP-RP & RUWE & $P$ & assoc & age & teff & rstar & mstar & rcr & qual & Nsectors \\
%\midrule
%\endhead
\startdata
TIC & $T$ & $d$ & $\bprp$ & RUWE & $P$ & Assoc & Age & $T_{\rm eff}$ & $R_\star$ & $M_\star$ & $R$$_{\rm cr}$ & Quality & $N_{\rm sector}$ \\
-- & mag & pc & mag & -- & hr & -- & Myr & K & $R_\odot$ & $M_\odot$ & $R_\star$ & -- & -- \\
\hline
\input{20230613_LGB_RJ_CPV_TABLE_selfnapplied_rounded_data}
\enddata
\tablecomments{This table includes \ngoods\ CQVs (\texttt{Quality} flag true), and \nmaybes\ candidate CQVs (\texttt{Quality} flag false).  The machine-readable version, available online, includes additional columns for the Gaia DR2 source identifiers and parameter uncertainties.  The age uncertainties are typically $\approx \pm 10\%$,
	but can be asymmetric.  The median temperature, radius, and mass uncertainties are $\pm XXX$\,K, $\pm XX$\%, and $\pm YY$\% respectively.
	$N_{\rm sector}$ denotes the number of TESS sectors for which data are expected to be acquired between July~2018 and Oct~2024.   Association names and provenance follow conventions adopted by \citet{2018ApJ...856...23G}: 
	ABDMG: AB~Doradus moving group \citep{2015MNRAS.454..593B}.
	ARG: Argus \citep{2019ApJ...870...27Z}.
	APER: $\alpha$~Persei open cluster \citep{2023AJ....166...14B}.
	BPMG: $\beta$~Pic moving group \citep{2015MNRAS.454..593B}.
  CARN: Carina Near moving group \citep{2006ApJ...649L.115Z}.
	COL: Columba \citep{2015MNRAS.454..593B}.
	EPSC: $\epsilon$ Chamaeleontis \citep{2013MNRAS.435.1325M}.
	LCC: Lower Centaurus Crux \citep{2016MNRAS.461..794P}.
	PERI: Pisces-Eridani \citep{2019AJ....158...77C}.
	PLE: Pleiades \citep{2015ApJ...813..108D}.
	TAU: Taurus \citep{1995ApJS..101..117K}.
	THA: Tucana-Horologium assocation \citep{2015MNRAS.454..593B}.
	TWA: TW Hydrae assocation \citep{2015MNRAS.454..593B}.
	UCL: Upper Centaurus Lupus \citep{2016MNRAS.461..794P}.
	USCO: Upper Scorpius \citep{2016MNRAS.461..794P}.
}
\end{deluxetable}
\clearpage

>>>>

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f3.pdf}
    \vspace{-0.3cm}
		\caption{
      {\bf CQVs found in the TESS 2-minute data.}
      Phased TESS light curves over one month are shown for \ngoods\
      CQVs.  Gray are raw 2-minute data; black bins to 300 points per
      cycle.  Each panel is labeled by the TIC identifier, the TESS
      sector number, and the period in hours.  Objects are ordered
      such that sources with the most TESS data available are on top
      (see Section~\ref{sec:catalog}).
		}
		\label{fig:cqvs}
	\end{center}
\end{figure*}

@@

%FIXME it'd be useful to get an RV for this one...
The upper age limit for CQVs may even pass 200\,Myr, based on the
candidate membership of TIC~294328887 in the Carina Near moving group
\citep{2006ApJ...649L.115Z}.  This group's $200 \pm 50$\,Myr age
primarily rests on the lithium sequence of its G-dwarfs
\citep{2006ApJ...649L.115Z}, which shows a coeval population of stars
older than the Pleiades and younger than the 400\,Myr Ursa Major
moving group.  The formal BANYAN-$\Sigma$ membership probability is
however somewhat low, perhaps due to a

It would interesting to perform independent tests of
the group's age, for instance using gyrochronology of F, G or K
members, in order to verify this .

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f4.pdf}
		\vspace{-0.6cm}
		\caption{
		{\bf CQVs keep their periods but change their shapes.}
      32 of our \ngoods\ CQVs from Figure~\ref{fig:cqvs} had
      120-second cadence TESS data available for a baseline of at
      least two years; the 27 brightest are shown here.  Each panel
      shows one sector of TESS data, and is phased to its deepest
      minimum in flux.  Each panel's title shows the TIC identifier
      and approximate period in hours.  Text insets show the TESS
      sector numbers, which generally span two years, or at least
      1{,}000 cycles.  The vertical scale is fixed across sectors to
      clarify shape changes.
      Gray circles are raw 2-minute data; colored circles bin to 300
      points per cycle. 
		}
		\label{fig:evoln}
	\end{center}
\end{figure*}

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f5.pdf}
		\vspace{-0.35cm}
		\caption{
      {\bf LP 12-502 (TIC~402980664) light curve}, where each time
      chunk represents one TESS orbit.  Data were acquired in Sectors
      18-19, 25-26, 53, and 57-58.  Flares are drawn in gray.  The red
      vertical lines highlight apparently instantaneous state changes
      in the shape of the dip pattern.  The light curve is binned to
      15-minute intervals so that there are 96 points per day, and each
      point is connected by a line.  
      Data gaps longer than 15-minutes are not
      interpolated; if data are missing, nothing is plotted.
		}
		\label{fig:lplc}
	\end{center}
\end{figure*}

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f6.pdf}
		\vspace{-0.6cm}
		\caption{
      {\bf Evolution of LP 12-502} ($P$=18.6\,h) at fixed period and
      epoch over three years.  Each panel shows one (stacked) TESS
      orbit; small text denotes relative cycle number.  There are 200
      binned black points per cycle.  The TESS pointing law dictates
      time gaps; larger gaps tend to yield larger shape changes.  The
      dips usually evolve over tens to hundreds of cycles.  However
      cycles 1233-1264 show a dip that ``switched'' from a depth and
      duration of 3\% and 3\,hr to 0.3\% and 1\,hr over less than one
      cycle (cf.~Figure~\ref{fig:lplc}).
		}
		\label{fig:lp}
	\end{center}
\end{figure*}

@@

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=0.9\textwidth]{f7.pdf}
		\vspace{-0.1cm}
		\caption{
      {\bf Properties of CQVs identified by our search}.
      The \ngoods\ bonafide CQVs in Table~\ref{tab:thetable} are 
      the dark blue circles; \nmaybes\ candidate CQVs are light blue circles.
      The top panels show the \nstarssearched\ target stars with 120-second cadence TESS data
      as the shaded gray background;
      darker regions correspond to a larger relative number of searched stars.
      The lower-left panel compares the rotation--color distribution of CQVs against the rotation
      periods of K- and M-dwarfs in the Pleiades from \citet{2016AJ....152..114R}.
      The lower-middle panel plots the derived corotation radii
      $R_{\rm cr} = (GM/\Omega^2)^{1/3}$ in units of stellar radii against the measured
      CQV periods, in units of hours.
		}
		\label{fig:catalogscatter}
	\end{center}
\end{figure*}

@@

\begin{enumerate}[leftmargin=*]
	%
  \item CQVs are quasiperiodic.  The mean periods 
    remain fixed over the $>$1{,}000-cycle baseline of available
    observations; but the light curve shapes always evolve (Figure~\ref{fig:evoln}).
  %
  \item The same CQV can show dips with similar but clearly
    distinct periods.  LP 12-502, for instance, showed dips with four
    distinct periods within $\pm 0.3\%$ of its fundamental period,
    sometimes simultaneously, and each lasting for up to 50 cycles
    (Figure~\ref{fig:lpriver0}).
  %
    \item CQVs evolve over timescales that are both secular
    ($>$100\,cycles) and impulsive ($<$1 cycle).  ``State-switches''
    can cause dips to collapse instantaneously, and are often (but not
    exclusively) linked with observed optical flares.  Dip growth
    however seems to happen over durations of at least ten cycles, and
    slow dip decay can also occur.
  %
    \item The rate of dip evolution can be used to place a model-dependent
    constraint on amount of material that is either being accreted or ejected
    during the state changes (Section~\ref{subsec:massflux}).
    Order of magnitude estimates require at least an asteroid belt's worth
    of dust ($10^{-4}$\,$M_\oplus$) over $10^8$ years, or 100 times more 
    material if the occulting matter is gas.
  %
  \item The CQV phenomenon persists for $\gtrsim$150\,Myr, based on
    the existence of multiple CQVs in AB Dor, the Pleiades, and
    Psc-Eri (Section~\ref{subsec:starprops}).  It may even extend to
    200\,Myr, based on the one CQV we found in the Carina Near
    moving group (TIC~294328887; $\approx$200\,Myr).  The lack of
    detected CQVs in the Hyades and Praesepe suggests that the
    lifetime of the phenomenon is likely limited to the first few
    hundred million years.
	%
    \item The duty cycle for CQVs seems to be $\approx 75$\%, based on
    the fraction of bonafide CQVs that turned on or turned off during
    TESS re-observations, two years after their initial observation.
  %
    \item While the majority of detected CQVs have masses of
    0.1-0.5\,M$_\odot$, the upper mass limit for the CQV phenomenon
    might extend into the K-dwarf regime ($<$$0.8$\,M$_\odot$).
    TIC~405754448 and TIC~405910546, with masses of $\approx$0.8 and
    $\approx$0.6 solar masses, are the two massive outliers
    (Section~\ref{subsec:extreme}), and we encourage additional
    scrutiny of these objects in future work.
  %
    \item The closest CQVs to the Sun are at distances of 15-20\,pc, and
    the brightest CQVs have $V$$\approx$12 ($J$$\approx$7.5).  This
    conclusion is an extrapolation from the CQV sample derived in this work
    being $\gtrsim$80\% complete within 20\,pc.  Expanding our analysis of
    the TESS data to the full frame images would yield a truly
    volume-limited selection function, and would expand the CQV census
    by about a factor of two within 50\,pc, and by a factor of ten
    within 100\,pc.
\end{enumerate}

@@

\acknowledgments
LGB gratefully acknowledges support from the
Heising-Simons 51~Pegasi~b Fellowship.
While many conversations helped encourage this work,
a few thank-yous in particular (from LGB) 
are owed to G.~Laughlin, A.~Mann, and J.~Spake for clarifications and ideas.
This paper relies on data collected by the TESS mission, which are publicly available from the Mikulski Archive for Space Telescopes (MAST). Funding for the TESS mission is provided by NASA’s Science Mission Directorate.
TIC 402980664 in particular was observed at 120-second cadence thanks to the TESS Guest
Investigator programs G022252 (PI: J.~Schlieder; Sectors 18, 19, 25,
26) and G04168 (PI: R.~Jayaraman; Sector 53).

@@

\startlongtable
\begin{deluxetable}{rrrrrrlrllrrrr}
\tabletypesize{\footnotesize}
%\tabletypesize{\scriptsize}
\tablecaption{Complex quasiperiodic variables identified in the TESS
  2-minute data. {\bf For internal review, full versions are available
  here
  \url{https://www.dropbox.com/scl/fo/twn4s9ckbevf75jqhtoy1/h?rlkey=t5cn8cx2uoc2ptdm9e570kp1b&dl=0}} \label{tab:thetable}}
%\toprule
%TIC & $T$ & $d$ & BP-RP & RUWE & $P$ & assoc & age & teff & rstar & mstar & rcr & qual & Nsectors \\
%\midrule
%\endhead
\startdata
TIC & $T$ & $d$ & $\bprp$ & RUWE & $P$ & Assoc & Age & $T_{\rm eff}$ & $R_\star$ & $M_\star$ & $R$$_{\rm cr}$ & Quality & $N_{\rm sector}$ \\
-- & mag & pc & mag & -- & hr & -- & Myr & K & $R_\odot$ & $M_\odot$ & $R_\star$ & -- & -- \\
\hline
\input{20230613_LGB_RJ_CPV_TABLE_selfnapplied_rounded_data}
\enddata
\tablecomments{This table includes \ngoods\ CQVs (\texttt{Quality} flag true), and \nmaybes\ candidate CQVs (\texttt{Quality} flag false).  The machine-readable version, available online, includes additional columns for the Gaia DR2 source identifiers and parameter uncertainties.  The age uncertainties are typically $\approx \pm 10\%$,
	but can be asymmetric.  The median temperature, radius, and mass uncertainties are $\pm XXX$\,K, $\pm XX$\%, and $\pm YY$\% respectively.
	$N_{\rm sector}$ denotes the number of TESS sectors for which data are expected to be acquired between July~2018 and Oct~2024.   Association names and provenance follow conventions adopted by \citet{2018ApJ...856...23G}: 
	ABDMG: AB~Doradus moving group \citep{2015MNRAS.454..593B}.
	ARG: Argus \citep{2019ApJ...870...27Z}.
	APER: $\alpha$~Persei open cluster \citep{2023AJ....166...14B}.
	BPMG: $\beta$~Pic moving group \citep{2015MNRAS.454..593B}.
  CARN: Carina Near moving group \citep{2006ApJ...649L.115Z}.
	COL: Columba \citep{2015MNRAS.454..593B}.
	EPSC: $\epsilon$ Chamaeleontis \citep{2013MNRAS.435.1325M}.
	LCC: Lower Centaurus Crux \citep{2016MNRAS.461..794P}.
	PERI: Pisces-Eridani \citep{2019AJ....158...77C}.
	PLE: Pleiades \citep{2015ApJ...813..108D}.
	TAU: Taurus \citep{1995ApJS..101..117K}.
	THA: Tucana-Horologium assocation \citep{2015MNRAS.454..593B}.
	TWA: TW Hydrae assocation \citep{2015MNRAS.454..593B}.
	UCL: Upper Centaurus Lupus \citep{2016MNRAS.461..794P}.
	USCO: Upper Scorpius \citep{2016MNRAS.461..794P}.
}
\end{deluxetable}
\clearpage
